Correlation between photons emitted in diferent channels from an atom inside a leaky driven cavity

### 2019-12-20 newpaper.tex
Version quase final, falta aprobacion de Guillermo
Version para mandar a EPJ D

Spontaneous emission has been studied since the early XX century
\cite{10.2307/94746, 1917PhyZ...18..121E}. One its main
characteristics is that the exact time of emission, as well as the
direction in which the photon is emitted, is random, which difficults
its measurement. On the other hand, photons that leak a cavity mirror
are easier to measure because we know their direction \cite{326305,
  doi:10.1063/1.113345}. The randomness of the direction in which the
photons are emitted by atoms in free space difficults photon counting
experiments. To sort out this difficulty one could think of a physical
arrangement in which a two-level atom is coupled to a driven cavity,
as it is shown in Fig. \ref{asa}. In this physical arrangement, we
have two emission channels: spontaneous emissions of the atom
--photons that are emitted in modes that are not cavity modes and
therefore escape the cavity to its side-- and cavity emissions
--photons that leak through one of the cavity mirrors. In the case
that a correlation exists between the two channels, we can count the
number of photons that leak the cavity through its mirror, which is
easy to measure, to find information about the number of photons
emitted into the other modes (spontaneous emission), which are
difficult to measure.

@@

We use quantum trajectories to calculate the joint probability of the
number of spontaneus emissions and the number of cavity emission. A
quantum trajectory describes a continuously monitored quantum system
\cite{Carmichael1993Open}. In our case each trajectory gives the
number of photons emitted in each channel. Simulating a large number
of trajectories and registering the number of photon emissions into
the two channels, we obtain its joint probability distribution and
their correlation. The correlation varies with the cavity linewidth
and the atom-cavity coupling. For a very bad cavity we recover the
free space result: there is no correlation between the two channels.

@@

\ENDIF
\ELSE
\STATE $\ket{ \phi(t + \delta t)} = \frac{\ket{\phi^1 (t+ \delta t)}}{\sqrt{1-\delta p}}$
\STATE $t \ += \delta t$
\ENDIF
\ENDFOR
\RETURN  $N_{CA}$
\end{algorithmic}
\end{algorithm}
We compare the collapse probability of Eq.~\eqref{probcol} with a
random number $\epsilon$, uniformly distributed between zero and one.
In the case the condition $\delta p > \epsilon$ is fulfilled we
simulate that a quantum jump has happened, by acting with one of the
collapse operators $a$ or $\sigma_-$ and renormalizating the obtained
state. The operator that will act on the state will be given according
to the relative probabilities of Eqs.~\eqref{relprob}.
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.pdf}
\caption{Probability distribution of spontaneous emissions for
  different number of cavity photons. Parameters $\mathcal{E}  =
  \kappa =  g = 1\gamma$, $100,000$ quantum trajectories. } \label{probdiss}
\end{figure}
If $\delta p > \epsilon$ is not fulfilled we just renormalize the
state, because using a non-hermitian Hamiltonian will have as a
consequence that the probability won't be conserved
\cite{Sakurai:1167961}. For the next step we repeat the procedure
using the new state. One quantum trajectory of total time $t_f$ is
obtained after repeating the previous procedure $t_f/\delta t$ steps.
For each trajectory we count the number of quantum jumps, the jumps
corresponding to operator $a$ reduces the number of photons inside the
cavity and corresponds to a photon that leak the cavity. The jumps
corresponding to operator $\sigma_-$, send the atom back to the ground
state and corresponds to a spontaneous emission. The procedure to
register emissions in the two different channels is summarized in
algorithm. \ref{alg1}.

@@

\section{Counting photons emitted in different channels}\label{sc:distributions}
In this section we calculate the probability distribution of the
number of spontaneous emissions $n_s$, under two different conditions.
First, we study the distribution probability of $n_s$ when we consider
the set of quantum trajectories where the number of leaked photons
from the cavity is $n_c$. This simulate the physical situation where
we count the number of spontaneous photons until the moment where the
number of leaked photons from the cavity is $n_c$. It is impossible to
know in advance how long it takes to the cavity to leak $n_c$ photons,
thus the duration of each quantum trajectory is not fixed. Second, we
fix the duration of each quantum trajectory and count the number of
spontaneous emissions $n_s$, and cavity emissions $n_c$. From the set
of quantum trajectories evolved a fixed time, we calculate the joint
probability $p(n_s,n_c)$, of finding $n_s$ spontaneous emissions and
$n_c$ cavity emissions. Using $p(n_s,n_c)$ we calculate the
conditional probability $p(n_s|n_c)$, and the correlations between the
two random variables. This correlation is a measure of how much
information we can obtain about $n_s$ if we know $n_c$.

@@

\subsection{Quantum trajectories with fixed number of photons leaked from the cavity}
We calculate the distribution probability of the number of spontaneous
emission when a fixed number of photons leaked the cavity. We assume
that the mean number of photons inside the cavity is $1/9$. The
results are shown in Fig.~\ref{probdiss}. As the number of cavity
emissions increases, the mean number of spontaneous emissions
increases. This result is expected, because the number of cavity
emissions is correlated with the time we are counting spontaneous
emission from the atom: the larger the number of cavity
emissions the larger the time we count spontaneous emission,
which implies a larger mean number of emitted photons. Also, as shown
in Fig~\ref{graph}, the probability distribution is wider as the
number of leaked photons increases, this implies less accuracy
guessing the number of spontaneous emission. If we know the cavity
drive, this results does not imply that we can learn something extra
about the number of spontaneous emissions knowing the number of
photons leaked from the cavity; it is clear that when we increase the
time window where we count photons, both, the number of cavity emissions
and the number of spontaneous emissions, increases.

@@

The results are shown in Fig. \ref{corrxy}. We observe a maximum
correlation, close to $0.1$, for $\kappa/\gamma\approx 0.05$. From
there the correlation decreases as $\kappa$ increases. The figure is
consistent with the statement that the correlation goes to zero as
$\kappa\rightarrow\infty$. With values for $\kappa/\gamma$ lower than
$0.05$, we notice that the correlations begins to decrease. The reason
is that, as the cavity linewidth decreases, the number of cavity
emissions decreases. In the limit where $\kappa=0$ the two random
variables becomes independent, because the output of $n_c$ is zero
with probability one. In this case there is no correlation between the
two random variables. Fig.~\ref{probdisult}(a), where the change in
$p(n_s|n_c)$ is larger, corresponds to a correlation around $0.1$;
Fig.~\ref{probdisult}(b), where the change is small, corresponds to a
correlation around $0.01$.

@@

\section{Conclusions}\label{sc:conclusions}
Atomic emission in cavity and non-cavity modes of an atom inside a
leaky cavity are correlated. The correlations are produced because the
cavity breaks the markovian behavior of atomic emission. The presence
of correlation between photons measured in different directions is a
signature of non-markovian behavior. This correlation allows us to
improve the knowledge we know about the probability distribution of
atomic spontaneous emission if we know the number of photons that leak
the cavity. In the system configuration that we studied the maximum
correlation is small. Nevertheless, we think that the results
encourage to explore different configurations where the correlation
can be larger.

>>>>

Spontaneous emission has been studied since the early XX century
\cite{10.2307/94746, 1917PhyZ...18..121E}. One of its main
characteristics is that the exact time of emission, as well as the
direction in which the photon is emitted, is random, which difficults
its measurement. On the other hand, photons that leak a cavity mirror
are easier to measure because we know their direction \cite{326305,
  doi:10.1063/1.113345}. The randomness of the direction in which the
photons are emitted by atoms in free space difficults photon counting
experiments. To sort out this difficulty one could think of a physical
arrangement in which a two-level atom is coupled to a driven cavity,
as it is shown in Fig. \ref{asa}. In this physical arrangement, we
have two emission channels: spontaneous emissions of the atom
--photons that are emitted in modes that are not cavity modes and
therefore escape the cavity to its side-- and cavity emissions
--photons that leak through one of the cavity mirrors. In the case
that a correlation exists between the two channels, we can count the
number of photons that leak the cavity through its mirror, which is
easy to measure, to find information about the number of photons
emitted into the other modes (spontaneous emission), which are
difficult to measure.

@@

We use quantum trajectories to calculate the joint probability of the
number of spontaneus emissions and the number of cavity emissions. A
quantum trajectory describes a continuously monitored quantum system
\cite{Carmichael1993Open}. In our case each trajectory gives the
number of photons emitted in each channel. Simulating a large number
of trajectories and registering the number of photon emissions into
the two channels, we obtain its joint probability distribution and
their correlation. The correlation varies with the cavity linewidth
and the atom-cavity coupling. For a very bad cavity we recover the
free space result: there is no correlation between the two channels.

@@

\ENDIF
\ELSE
\STATE $\ket{ \phi(t + \delta t)} = \frac{\ket{\phi^1 (t+ \delta t)}}{\sqrt{1-\delta p}}$
\STATE $t \ += \delta t$
\ENDIF
\ENDFOR
\RETURN  $N_{CA}$
\end{algorithmic}
\end{algorithm}
We compare the collapse probability of Eq.~\eqref{probcol} with a
random number $\epsilon$, uniformly distributed between zero and one.
In the case the condition $\delta p > \epsilon$ is fulfilled we
simulate that a quantum jump has happened, by acting with one of the
collapse operators $a$ or $\sigma_-$ and renormalizating the obtained
state. The operator that will act on the state will be given according
to the relative probabilities of Eqs.~\eqref{relprob}.
\begin{figure}[!t] 
\centering
\includegraphics[scale = 0.5]{distributioneng.pdf}
\caption{Probability distribution of spontaneous emissions for
  different number of cavity photons. Parameters: $\expval{n} \approx 1$, $\kappa = g = 0.1\gamma$. } \label{probdiss}
\end{figure}
If $\delta p > \epsilon$ is not fulfilled we just renormalize the
state, because using a non-hermitian Hamiltonian will have as a
consequence that the probability won't be conserved
\cite{Sakurai:1167961}. For the next step we repeat the procedure
using the new state. One quantum trajectory of total time $t_f$ is
obtained after repeating the previous procedure $t_f/\delta t$ steps.
For each trajectory we count the number of quantum jumps, the jumps
corresponding to operator $a$ reduces the number of photons inside the
cavity and corresponds to a photon that leak the cavity. The jumps
corresponding to operator $\sigma_-$, send the atom back to the ground
state and corresponds to a spontaneous emission. The procedure to
register emissions in the two different channels is summarized in
algorithm. \ref{alg1}.

@@

\section{Counting photons emitted in different channels}\label{sc:distributions}
In this section we calculate the probability distribution of the
number of spontaneous emissions $n_s$, under two different conditions.
First, we study the distribution probability of $n_s$ when we consider
the set of quantum trajectories where the number of leaked photons
from the cavity is $n_c$. This simulates the physical situation where
we count the number of spontaneous photons until the moment where the
number of leaked photons from the cavity is $n_c$. It is impossible to
know in advance how long it takes to the cavity to leak $n_c$ photons,
thus the duration of each quantum trajectory is not fixed. Second, we
fix the duration of each quantum trajectory and count the number of
spontaneous emissions $n_s$, and cavity emissions $n_c$. From the set
of quantum trajectories evolved a fixed time, we calculate the joint
probability $p(n_s,n_c)$, of finding $n_s$ spontaneous emissions and
$n_c$ cavity emissions. Using $p(n_s,n_c)$ we calculate the
conditional probability $p(n_s|n_c)$, and the correlations between the
two random variables. This correlation is a measure of how much
information we can obtain about $n_s$ if we know $n_c$.

@@

\subsection{Quantum trajectories with fixed number of photons leaked from the cavity}
We calculate the distribution probability of the number of spontaneous
emission when a fixed number of photons leaked the cavity. We assume
that the mean number of photons inside the cavity is $1$. The
results are shown in Fig.~\ref{probdiss}. As the number of cavity
emissions increases, the mean number of spontaneous emissions
increases. This result is expected, because the number of cavity
emissions is correlated with the time we are counting spontaneous
emission from the atom: the larger the number of cavity
emissions the larger the time we count spontaneous emission,
which implies a larger mean number of emitted photons. Also, as shown
in Fig~\ref{graph}, the probability distribution is wider as the
number of leaked photons increases, this implies less accuracy
guessing the number of spontaneous emission. If we know the cavity
drive, this results does not imply that we can learn something extra
about the number of spontaneous emissions knowing the number of
photons leaked from the cavity; it is clear that when we increase the
time window where we count photons, both, the number of cavity emissions
and the number of spontaneous emissions, increases.

@@

The results are shown in Fig. \ref{corrxy}. We observe a maximum
correlation, close to $0.1$, for $\kappa/\gamma\approx 0.05$. From
there the correlation decreases as $\kappa$ increases. The figure is
consistent with the statement that the correlation goes to zero as
$\kappa\rightarrow\infty$. With values for $\kappa/\gamma$ lower than
$0.05$, we notice that the correlations begins to decrease. The reason
is that, as the cavity linewidth decreases, the number of cavity
emissions decreases. In the limit where $\kappa=0$ the two random
variables become independent, because the output of $n_c$ is zero
with probability one. In this case there is no correlation between the
two random variables. Fig.~\ref{probdisult}(a), where the change in
$p(n_s|n_c)$ is larger, corresponds to a correlation around $0.1$;
Fig.~\ref{probdisult}(b), where the change is small, corresponds to a
correlation around $0.01$.

@@

\section{Conclusions}\label{sc:conclusions}
Atomic emissions in cavity and non-cavity modes of an atom inside a
leaky cavity are correlated. The correlations are produced because the
cavity breaks the markovian behavior of atomic emission. The presence
of correlation between photons measured in different directions is a
signature of non-markovian behavior. This correlation allows us to
improve the knowledge we know about the probability distribution of
atomic spontaneous emission if we know the number of photons that leak
the cavity. In the system configuration that we studied the maximum
correlation is small. Nevertheless, we think that the results
encourage to explore different configurations where the correlation
can be larger.
